### 2017-08-22 thesis/results/results.tex
Final commit before submission of this thesis
Added marking_response to the thesis directory temporarily

\chapter{Obesity associated genetic signatures and cancer}
\label{cha:obesity_genetic_signatures_and_cancer}

The obesity associated genetic signatures are central to this project, as a means to clarify the relationship between \gls{bmi} and tumour biology.
In this chapter, the previously identified obesity associated genetic signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} are examined in turn to judge the agreement of these signatures with the patient \gls{bmi} and \gls{bmi} status, as presented in their results.
After this, novel obesity associated genetic signatures are identified in the Creighton \textit{et al.} (CR) data set and was compared with the obesity associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014}  studies.
Lastly, the presence of common genes or pathways associated with obesity in multiple types of cancer is explored using the data sets from \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}).

\section{Obesity associated genetic signature from the\\\citet{Creighton2012} study}
\label{sec:creighton_obesity_metagene}

An obesity metagene was created using the obesity associated genetic signature from the \citet{Creighton2012} study.
There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the more popular \gls{rma} method  was used to  normalise the CR data set \citep{Irizarry2003}.
The obesity metagene was created in the standardised \gls{rma} normalised CR data set, and  was plotted above a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).

\begin{figure}[tb!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
	\caption[Obesity metagene from the \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity associated genes in the CR data.
	Level of expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the CR data.
	The obesity associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the hierarchical clustering of the genes is shown on the left of the heatmap.
	For clarity, the sign of the metagene scores were reversed in order to match the results from the \citet{Creighton2012} study.}
	\label{fig:crmetaheat}
\end{figure}

As shown in \cref{fig:crmetaheat}, a high obesity associated metagene score  reflects low expression in majority of the genes in the signature, and in contrast, a low obesity associated metagene score reflects high expression in the majority of the genes in the signature.
This was consistent with the reported property of the obesity associated genetic signature by \citet{Creighton2012} (see \cref{ssub:creighton_study}).
To provide further evidence that the obesity metagenes were in fact associated with both the \gls{bmi} status and the \gls{bmi} value of the patients, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).

@@

\cref{fig:crmetaboxplot} clearly showed that the obesity metagene from the \citet{Creighton2012} study significantly associated with the obese group of patients, as well as patient \gls{bmi} value.
It should be noted here that the obesity metagene significantly associated with the samples from the patients that were obese, but not with the sample from the patients that were overweight.
This was due to the fact that the obesity associated genetic signature was originally identified from the comparison of the patients that were obese with the patients that were not obese, and therefore the obesity metagene scores were significant with the obese group, but not with the overweight group.
Another thing to note here was that even though  the regression line in the scatter plot showed statistically significant association with the patient \gls{bmi}, patient \gls{bmi} values seem to be randomly dispersed across the obesity metagene scores.
This suggests that perhaps the obesity metagene from the \citet{Creighton2012} study may not be associated with the patient \gls{bmi} as strongly as reported.
\\

@@

There could be several reasons for the apparent lack of association of the obesity metagene with patient \gls{bmi} in most of the \gls{icgc} data.
First, the transformation matrix was derived from the CR microarray data, but the \gls{icgc} cancer data were generated via \gls{rnaseq}.
Though the log$_{10}$-normalisation and standardisation of the data was the most appropriate adjustments to be made to the \gls{rnaseq} data, these adjustments were not equivalent to the \gls{rma} normalisation method that was used on the microarray data.
Secondly, none of the \gls{icgc} cancer data in this project originated from breast tumours as in the CR data.
Since the obesity associated genetic signature was identified in the breast cancer data, the signature may be specific to breast cancer and may not be generalisable to other cancer types.
\\

@@

The Creighton \textit{et al.} obesity metagene managed to reflect the overall gene expression of the samples in the \gls{nzbc}  data (\cref{fig:crmetaprint}).
However, as with the \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with patient \gls{bmi} or \gls{bmi} status (\cref{fig:crmetaprint}).
These results confirmed that the lack of association of the obesity associated genetic signature from the Creighton \textit{et al.} study was not due to the technology in which the data was gathered (microarray or \gls{rnaseq}), nor the cancer type in which the genetic signature was derived from.

@@

Taken together, these results suggest that the obesity associated genetic signature identified in the study conducted by \citet{Creighton2012} correlated with patient \gls{bmi} and \gls{bmi} status only in the CR data set.
The lack of association with patient \gls{bmi} in other cancer data sets was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor the \gls{nzbc} microarray data showed significant association of the obesity metagene with the patient \gls{bmi}.
Furthermore, the obesity associated genetic signature was not dependent on the cancer type in which it was generated from, since the obesity metagene did not show significant association in either the \gls{icgc} cancer data or in the \gls{nzbc} data.

One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable (investigated in \cref{sec:creighton_obesity_metagene_new}).
Another reason for this apparent lack of association could be that the obesity associated genetic signature was too specific to the CR data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the patient cohort that was profiled in the \citet{Creighton2012} publication.

\section{Obesity associated genetic signature from the\\\citet{Fuentes-Mattei2014} study}
\label{sec:fm_obesity_metagene}

The obesity associated genetic signature from the \citet{Creighton2012} study was not associated with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets, so the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether this obesity metagene was able to significantly associate with the patient \gls{bmi} and \gls{bmi} status across different  data sets.
Since the FM data set did not have patient \gls{bmi} information, the FM obesity metagene was not able to be compared with the patient \gls{bmi} or \gls{bmi} status in the original FM data.
Nevertheless, the transformation matrix was still generated in the FM data and applied first to the microarray data (CR and \gls{nzbc} data sets) then to the \gls{icgc} data sets to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in these data sets.

@@

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_fm_meta}\\
	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_fm_meta}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the FM obesity associated metagene with the sample gene expression, patient  \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:fmmetacris}
\end{figure}

Next, the transformation matrix was applied to the \gls{icgc} cancer data and the resulting FM metagene was compared with the sample gene expression, patient \gls{bmi} and \gls{bmi} status in each of the \gls{icgc} cancer data.
As evident in \cref{fig:fmmetaicgc,sec:rest_of_the_fm_icgc_cancer_heatmap_results}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity associated genetic signature.
As with all the results in this chapter so far, the FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in the \gls{blca} data set (\cref{fig:fmmetaicgc}; \cref{sec:rest_of_the_fm_icgc_cancer_heatmap_results}).
The FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value \cref{fig:fmmetaicgc}.
On the contrary to the association of the metagene with the patient \gls{bmi} status, the FM obesity metagene was not associated with the patient \gls{bmi}.
These results suggested that the samples from the patients that were overweight in the \gls{blca} data set had similar biological properties as the samples taken from the patients that were obese in the FM data set.
However, due to the fact that the FM obesity metagene lacked association with the patient \gls{bmi} in the \gls{blca} data set, and that the metagene did not show any significant association in any of the other \gls{icgc} cancer types, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group was truly reflective of the quality of the FM metagene or observed by chance.

@@

These results showed that the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was not generalisable to other cancer data sets, similar to the obesity signature identified by Creighton \textit{et al.}.
This meant that both the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} obesity metagenes may have been too specific to the original data set in which the signatures were identified in.
Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).

\section{Novel obesity associated genetic signatures from the \citet{Creighton2012} data set}
\label{sec:creighton_obesity_metagene_new}

\subsection{Identification of novel obesity associated genetic signatures}
\label{sub:identification_of_obesity_associated_genetic_signatures}

Both the obesity associated genetic signatures identified from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies may not have been truly associated with patient \gls{bmi}, but with another clinical variable.
To investigate this possibility, novel obesity associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
The FM data set was not used to get the obesity associated genetic signatures, as no \gls{bmi} information was available for the patients in the FM data.

Firstly, an attempt was made to replicate the Creighton \textit{et al.} obesity associated genetic signature from the original CR data.
\citet{Creighton2012} originally found their obesity associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, Creighton \textit{et al.} selected the 799 statistically significant \glspl{deg} (p \textless{} 0.01) with log$_2$-fold change greater than 1.2.
Therefore, gene expression analysis  between the samples from obese and non-obese patients in the \gls{rma} normalised CR data was carried out (described in \cref{sec:gene_expression_analysis}).
% Therefore, \glspl{deg}  were identified between the samples from obese and non-obese patients in the \gls{rma} normalised CR data, as described in \cref{sec:gene_expression_analysis}.
Without adjusting the p-value to account for multiple hypothesis testing, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively (\cref{tab:ob_deg_summary}).
After adjustments were made for multiple hypothesis testing, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggest that the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.

Furthermore, there were only 61 gene probes that were significantly differentially expressed at p \textless{} 0.05 with a log$_2$-\gls{fc} greater than 1.2.
From these observations, the log$_2$-fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (unadjusted p-value) for the identification of the significant probe sets.
Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity associated genetic signature, as this many genes were originally identified by Creighton \textit{et al.}.
This was to include as many probe set as possible for the novel genetic signatures to be comparable with the original Creighton \textit{et al.} obesity associated genetic signature.
These criteria were applied for the identification of other genetic signatures as well.

% To determine whether the identified obesity associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.

The above analysis was repeated with the residual data (\gls{rma} normalised CR data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
The clinical variables controlled were age, ethnicity, menopause status, tumour grade, hormone (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
In the residual data, 1104 gene probes were significantly differentially expressed with unadjusted p-value (p \textless{} 0.01; additional results shown in \cref{tab:ob_deg_summary}).
Again, the most significant 799 gene  probes were taken as the obesity associated genetic signature from this data.

In addition to the above two obesity associated genetic signatures, two more sets of genetic signatures were identified by taking the gene probes that were common between each of the above two signatures with the original obesity associated genetic signatures.
There were 239 common gene probes between the original Creighton \textit{et al.} obesity associated  genetic signature and the gene probes identified from the unadjusted CR data, and 168 common gene probes between the original signature and the gene probes from the residual CR data.
The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
\\

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
	\caption[Venn diagram of the \glspl{deg} identified from the CR data (all patients included)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data and the original obesity associated genetic signature from the \citet{Creighton2012} study }
	\label{fig:venn1}
\end{figure}

@@

With the Caucasian-only CR data set, obesity associated genetic signatures were identified as described above.
% ; first in the unadjusted data, then in the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering only the Caucasian samples), and lastly the common genes between each of these two signatures with the original obesity associated genetic signature were identified.
2129 and 1558 gene probes were significantly differentially expressed (unadjusted p-value \textless{} 0.01) in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively (\cref{tab:ob_deg_summary}).
As before, the most significant 799 gene probes were selected from these gene probes.
There were 148 and 92 common gene probes with the original Creighton \textit{et al.} obesity associated genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).

\begin{figure}[tb]
	\centering
	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
	\caption[Venn diagram of the \glspl{deg} identified from the CR data (Caucasian patients only)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data (Caucasian patients only) and the original obesity associated genetic signature from the \citet{Creighton2012} study}
	\label{fig:venn2}
\end{figure}

@@

\begin{figure}[tpb]
	\centering
	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from the CR data with one another.
	\gls{svd} was applied to the \gls{rma} normalised CR data to generate each of the eight metagenes.
	High and low correlation were represented as red and yellow, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.
	}
	\label{fig:cr_meta_cor}
\end{figure}

Since all of these obesity associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity associated genetic signature was searched for using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with patient \gls{bmi} or \gls{bmi} status in the \gls{nzbc} and \gls{icgc} cancer data sets (\cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}).
For simplicity, the abbreviations in \cref{tab:mg_abbrev} will be used to refer to the appropriate genetic signatures.

>>>>

\chapter{Obesity-associated genetic signatures and cancer}
\label{cha:obesity_genetic_signatures_and_cancer}

The obesity-associated genetic signatures are central to this project, as a means to clarify the relationship between \gls{bmi} and tumour biology.
In this chapter, the previously identified obesity-associated genetic signatures from the studies conducted by \citet{Creighton2012} and \citet{Fuentes-Mattei2014} are examined in turn to judge the agreement of these signatures with the patient \gls{bmi} and \gls{bmi} status, as presented in their results.
After this, novel obesity-associated genetic signatures were identified in the Creighton \textit{et al.} (CR) data set and was compared with the obesity-associated genetic signatures from the \citet{Creighton2012} and the \citet{Fuentes-Mattei2014}  studies.
Lastly, the presence of common genes or pathways associated with obesity in multiple types of cancer is explored using the data sets from \gls{icgc} (\citetalias{ICGC}; \citealp{Zhang2011}).

\section{Obesity-associated genetic signature from the\\\citet{Creighton2012} study}
\label{sec:creighton_obesity_metagene}

An obesity metagene was created using the obesity-associated genetic signature from the \citet{Creighton2012} study.
There was no description about the normalisation method used by Creighton \textit{et al.} when they first analysed their data, so the more popular \gls{rma} method  was used to  normalise the CR data set \citep{Irizarry2003}.
The obesity metagene was created in the standardised \gls{rma} normalised CR data set, and  was plotted above a heatmap with the sample gene expression  to check whether the metagene scores were in accordance with the overall gene expression of the samples (\cref{fig:crmetaheat}).

\begin{figure}[tb!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
	\caption[Obesity metagene from the \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity-associated genes in the CR data.
	Level of expression is represented in the top left histogram, where low and high gene expression are colour-coded with blue and red, respectively.
	Each row of the heatmap represents a gene from the obesity-associated genetic signature, and each column of the heatmap represents a sample from the CR data.
	The obesity-associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the hierarchical clustering of the genes is shown on the left of the heatmap.
	For clarity, the sign of the metagene scores were reversed in order to match the results from the \citet{Creighton2012} study.}
	\label{fig:crmetaheat}
\end{figure}

As shown in \cref{fig:crmetaheat}, a high obesity-associated metagene score  reflects low expression in majority of the genes in the signature, and in contrast, a low obesity-associated metagene score reflects high expression in the majority of the genes in the signature.
This was consistent with the reported property of the obesity-associated genetic signature by \citet{Creighton2012} (see \cref{ssub:creighton_study}).
To provide further evidence that the obesity metagenes were in fact associated with both the \gls{bmi} status and the \gls{bmi} value of the patients, a box plot and a scatter plot were created, respectively (\cref{fig:crmetaboxplot}).

@@

\cref{fig:crmetaboxplot} clearly showed that the obesity metagene from the \citet{Creighton2012} study significantly associated with the obese group of patients, as well as patient \gls{bmi} value.
It should be noted here that the obesity metagene significantly associated with the samples from the patients that were obese, but not with the sample from the patients that were overweight.
This was due to the fact that the obesity-associated genetic signature was originally identified from the comparison of the patients that were obese with the patients that were not obese, and therefore the obesity metagene scores were significant with the obese group, but not with the overweight group.
Another thing to note here was that even though  the regression line in the scatter plot showed statistically significant association with the patient \gls{bmi}, patient \gls{bmi} values seem to be randomly dispersed across the obesity metagene scores.
This suggests that perhaps the obesity metagene from the \citet{Creighton2012} study may not be associated with the patient \gls{bmi} as strongly as reported.
\\

@@

There could be several reasons for the apparent lack of association of the obesity metagene with patient \gls{bmi} in most of the \gls{icgc} data.
First, the transformation matrix was derived from the CR microarray data, but the \gls{icgc} cancer data were generated via \gls{rnaseq}.
Though the log$_{10}$-normalisation and standardisation of the data were the most appropriate adjustments to be made to the \gls{rnaseq} data, these adjustments were not equivalent to the \gls{rma}-normalisation method that was used on the microarray data.
Secondly, none of the \gls{icgc} cancer data in this project originated from breast tumours as in the CR data.
Since the obesity-associated genetic signature was identified in the breast cancer data, the signature may be specific to breast cancer and may not be generalisable to other cancer types.
\\

@@

The Creighton \textit{et al.} obesity metagene managed to reflect the overall gene expression of the samples in the \gls{nzbc}  data (\cref{fig:crmetaprint}).
However, as with the \gls{icgc} cancer data, the obesity metagene scores did not significantly associate with patient \gls{bmi} or \gls{bmi} status (\cref{fig:crmetaprint}).
These results confirmed that the lack of association of the obesity-associated genetic signature from the Creighton \textit{et al.} study was not due to the technology in which the data was gathered (microarray or \gls{rnaseq}), nor the cancer type in which the genetic signature was derived from.

@@

Taken together, these results suggest that the obesity-associated genetic signature identified in the study conducted by \citet{Creighton2012} correlated with patient \gls{bmi} and \gls{bmi} status only in the CR data set.
The lack of association with patient \gls{bmi} in other cancer data sets was not due to the type of technology platform in which the data was gathered, as neither the \gls{icgc} \gls{rnaseq} data nor the \gls{nzbc} microarray data showed significant association of the obesity metagene with the patient \gls{bmi}.
Furthermore, the obesity-associated genetic signature was not dependent on the cancer type in which it was generated from, since the obesity metagene did not show significant association in either the \gls{icgc} cancer data or in the \gls{nzbc} data.

One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable (investigated in \cref{sec:creighton_obesity_metagene_new}).
Another reason for this apparent lack of association could be that the obesity-associated genetic signature was too specific to the CR data and was not a broad obesity-associated genetic signature, but an obesity-associated signature that was specific to the patient cohort that was profiled in the \citet{Creighton2012} publication.

\vspace{-3mm}

\section{Obesity-associated genetic signature from the\\\citet{Fuentes-Mattei2014} study}
\label{sec:fm_obesity_metagene}

The obesity-associated genetic signature from the \citet{Creighton2012} study was not associated with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets, so the obesity-associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether this obesity metagene was able to significantly associate with the patient \gls{bmi} and \gls{bmi} status across different  data sets.
Since the FM data set did not have patient \gls{bmi} information, the FM obesity metagene was not able to be compared with the patient \gls{bmi} or \gls{bmi} status in the original FM data.
Nevertheless, the transformation matrix was still generated in the FM data and applied first to the microarray data (CR and \gls{nzbc} data sets) then to the \gls{icgc} data sets to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in these data sets.

@@

\begin{figure}[htp!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/cris_fm_meta}\\
	\includegraphics[page=4,width=0.45\linewidth]{results1/cris_fm_meta}
	\hfill
	\includegraphics[page=5,width=0.45\linewidth]{results1/cris_fm_meta}
	\caption[FM metagene in the \gls{nzbc} data]{Heatmap, scatter plot and box plot showing the association of the FM obesity-associated metagene with the sample gene expression, patient  \gls{bmi} and \gls{bmi} status, respectively, from the \gls{nzbc} data.
	Scales, p-values and $R^2$-value are as described in previous figures.}
	\label{fig:fmmetacris}
\end{figure}

Next, the transformation matrix was applied to the \gls{icgc} cancer data and the resulting FM metagene was compared with the sample gene expression, patient \gls{bmi} and \gls{bmi} status in each of the \gls{icgc} cancer data.
As evident in \cref{fig:fmmetaicgc,sec:rest_of_the_fm_icgc_cancer_heatmap_results}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity-associated genetic signature.
As with all the results in this chapter so far, the FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in the \gls{blca} data set (\cref{fig:fmmetaicgc}; \cref{sec:rest_of_the_fm_icgc_cancer_heatmap_results}).
The FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value \cref{fig:fmmetaicgc}.
On the contrary to the association of the metagene with the patient \gls{bmi} status, the FM obesity metagene was not associated with the patient \gls{bmi}.
These results suggested that the samples from the patients that were overweight in the \gls{blca} data set had similar biological properties as the samples taken from the patients that were obese in the FM data set.
However, due to the fact that the FM obesity metagene lacked association with the patient \gls{bmi} in the \gls{blca} data set, and that the metagene did not show any significant association in any of the other \gls{icgc} cancer types, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group was truly reflective of the quality of the FM metagene or observed by chance.

@@

These results showed that the obesity-associated genetic signature from the \citet{Fuentes-Mattei2014} study was not generalisable to other cancer data sets, similar to the obesity signature identified by Creighton \textit{et al.}.
This meant that both the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} obesity metagenes may have been too specific to the original data set in which the signatures were identified in.
Furthermore, there was a possibility that these obesity-associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).

\section{Novel obesity-associated genetic signatures from the \citet{Creighton2012} data set}
\label{sec:creighton_obesity_metagene_new}

\subsection{Identification of novel obesity-associated genetic signatures}
\label{sub:identification_of_obesity_associated_genetic_signatures}

Both the obesity-associated genetic signatures identified from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
One possible reason for this result could be that the obesity-associated genetic signatures from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies may not have been truly associated with patient \gls{bmi}, but with another clinical variable.
To investigate this possibility, novel obesity-associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
The FM data set was not used to get the obesity-associated genetic signatures, as no \gls{bmi} information was available for the patients in the FM data.

Firstly, an attempt was made to replicate the Creighton \textit{et al.} obesity-associated genetic signature from the original CR data.
\citet{Creighton2012} originally found their obesity-associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, Creighton \textit{et al.} selected the 799 statistically significant \glspl{deg} (p \textless{} 0.01) with log$_2$-fold change greater than 1.2.
Therefore, gene expression analysis  between the samples from obese and non-obese patients in the \gls{rma} normalised CR data was carried out (described in \cref{sec:gene_expression_analysis}).
% Therefore, \glspl{deg}  were identified between the samples from obese and non-obese patients in the \gls{rma} normalised CR data, as described in \cref{sec:gene_expression_analysis}.
Without adjusting the p-value to account for multiple hypothesis testing, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively (\cref{tab:ob_deg_summary}).
After adjustments were made for multiple hypothesis testing, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggest that the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.

Furthermore, there were only 61 gene probes that were significantly differentially expressed at p \textless{} 0.05 with a log$_2$-\gls{fc} greater than 1.2.
From these observations, the log$_2$-fold change of the gene probes were ignored and the threshold p-value was set to 0.01 (unadjusted p-value) for the identification of the significant probe sets.
Additionally, when there were more than 799 gene probes identified, only the most significant 799 gene probes were taken as the obesity-associated genetic signature, as this many genes were originally identified by Creighton \textit{et al.}.
This was to include as many probe set as possible for the novel genetic signatures to be comparable with the original Creighton \textit{et al.} obesity-associated genetic signature.
These criteria were applied for the identification of other genetic signatures as well.

% To determine whether the identified obesity-associated genetic signature was similar to the signature that Creighton \textit{et al.} had originally found, the identified \glspl{deg} were compared with the original genetic signature identified by Creighton \textit{et al.}.

The above analysis was repeated with the residual data (\gls{rma} normalised CR data that had been controlled for other clinical variables; see \cref{sub:residual_data_creation}).
The clinical variables controlled were age, ethnicity, menopause status, tumour grade, receptor (\gls{er}, \gls{pr}, and \gls{her2}) statuses  and \gls{ln} status.
In the residual data, 1104 gene probes were significantly differentially expressed with unadjusted p-value (p \textless{} 0.01; additional results shown in \cref{tab:ob_deg_summary}).
Again, the most significant 799 gene  probes were taken as the obesity-associated genetic signature from this data.

In addition to the above two obesity-associated genetic signatures, two more sets of genetic signatures were identified by taking the gene probes that were common between each of the above two signatures with the original obesity-associated genetic signatures.
There were 239 common gene probes between the original Creighton \textit{et al.} obesity-associated  genetic signature and the gene probes identified from the unadjusted CR data, and 168 common gene probes between the original signature and the gene probes from the residual CR data.
The genetic signatures identified were summarised as a Venn diagram, as shown in \cref{fig:venn1}.
\\

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.5\linewidth]{results1/deg_venn1}
	\caption[Venn diagram of the \glspl{deg} identified from the CR data (all patients included)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data and the original obesity-associated genetic signature from the \citet{Creighton2012} study }
	\label{fig:venn1}
\end{figure}

@@

With the Caucasian-only CR data set, obesity-associated genetic signatures were identified as described above.
% ; first in the unadjusted data, then in the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering only the Caucasian samples), and lastly the common genes between each of these two signatures with the original obesity-associated genetic signature were identified.
2129 and 1558 gene probes were significantly differentially expressed (unadjusted p-value \textless{} 0.01) in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively (\cref{tab:ob_deg_summary}).
As before, the most significant 799 gene probes were selected from these gene probes.
There were 148 and 92 common gene probes with the original Creighton \textit{et al.} obesity-associated genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).

\begin{figure}[tb]
	\centering
	\includegraphics[width=0.5\linewidth]{results1/deg_venn2}
	\caption[Venn diagram of the \glspl{deg} identified from the CR data (Caucasian patients only)]{Venn diagram showing the common gene probes between the signatures obtained from the unadjusted and the clinical variable-adjusted CR data (Caucasian patients only) and the original obesity-associated genetic signature from the \citet{Creighton2012} study}
	\label{fig:venn2}
\end{figure}

@@

\begin{figure}[tpb]
	\centering
	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity-associated metagenes from the CR data with one another.
	\gls{svd} was applied to the \gls{rma} normalised CR data to generate each of the eight metagenes.
	High and low correlation were represented as red and yellow, respectively, where the colours were matched with the values on the scale shown in the top left histogram.
	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.
	}
	\label{fig:cr_meta_cor}
\end{figure}

Since all of these obesity-associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity-associated genetic signature was searched for using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
All of these obesity-associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with patient \gls{bmi} or \gls{bmi} status in the \gls{nzbc} and \gls{icgc} cancer data sets (\cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}).
For simplicity, the abbreviations in \cref{tab:mg_abbrev} will be used to refer to the appropriate genetic signatures.

\begin{table}[tb]
	\centering
	\caption{Summary of the abbreviations used to refer to the different obesity-associated genetic signatures identified in the CR data}
	\label{tab:mg_abbrev}
	\begin{tabular}{lp{0.51\textwidth}c}
		\hline
		\hline
		Abbreviation & Definition & No. of gene probes\\
		\hline
		\rule{0pt}{2.25ex}Or      & Original obesity-associated genetic signature identified by \citet{Creighton2012}                       & 799\\
		\rule{0pt}{2.25ex}Cr      & Obesity-associated genetic signature identified from unadjusted CR data                                & 799 \\
		\rule{0pt}{2.25ex}CrOl    & Genes common between Or and Cr genetic signatures                                                       & 239\\
		\rule{0pt}{2.25ex}Res     & Obesity-associated genetic signature identified from clinical variable-adjusted CR data                & 799\\
		\rule{0pt}{2.25ex}ResOl   & Genes common between Or and Res genetic signatures                                                      & 168\\
		\rule{0pt}{2.25ex}Ca      & Obesity-associated genetic signature identified from unadjusted Caucasian-only CR data                 & 799\\
		\rule{0pt}{2.25ex}CaOl    & Genes common between Or and Ca genetic signatures                                                       & 148\\
		\rule{0pt}{2.25ex}CaRes   & Obesity-associated genetic signature identified from clinical variable-adjusted Caucasian-only CR data & 799\\
		\rule{0pt}{2.25ex}CaResOl & Genes common between Or and CaRes genetic signatures                                                    & 92\\
		\hline
		\hline
	\end{tabular}
\end{table}

\subsection{Novel obesity-associated signatures and patient \gls{bmi}/\gls{bmi} status}
\label{sub:_novel_obesity_associated_signatures_and_sample_bmi}
